Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-11 arxiv/proof_main.tex
more comment
edit

\subsection{The Bias and Variance Asymptotics}
Finally, in this subsection, we describe briefly the main ideas in the proofs of Lemmas \ref{lem_cov_shift} and \ref{lem_cov_derivative}.
%We first describe the proof of Theorem \ref{lem_cov_shift_informal}.
%We briefly describe the key ideas for proving Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative}.
We use the standard Stieltjes transform method (or the resolvent method) in random matrix theory \cite{bai2009spectral,tao2012topics,erdos2017dynamical}. Both Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative} require a study of the matrix inverse $(X_1^\top X_1+X_2^\top X_2)^{-1}$ for $X_1= Z_1\Sigma_1^{1/2}$ and $X_2= Z_2\Sigma_2^{1/2}$. Suppose $M=\Sig_1^{1/2} \Sig_2^{-1/2}$ has a singular value decomposition
\be\label{eigen2}
M= U\Lambda V^\top, \quad \Lambda=\text{diag}( \lambda_1, \ldots, \lambda_p).
\ee
Then using equation \eqref{eigen2}, we can write
\be\label{eigen2extra}(X_1^\top X_1+X_2^\top X_2)^{-1}= \Sigma_2^{-1/2}V\left(   \Lambda U^\top Z_1^\top Z_1 U\Lambda  + V^\top Z_2^\top Z_2V\right)^{-1}V^\top\Sigma_2^{-1/2}.\ee
We divide the discussion into the following parts.

@@

 Using Schur complement formula (see {equation} (\ref{resolvent2})), we have that %for $i \in \cal I_0$, $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\begin{align}
\frac{1}{{G_{ii} }}&=  - z - \frac{\lambda_i^2}{n} \sum_{\mu,\nu\in \mathcal I_1} z_{\mu i}z_{\nu i}G^{\left( i \right)}_{\mu\nu} - \frac{1}{n} \sum_{\mu,\nu\in \mathcal I_2} z_{\mu i}z_{\nu i}G^{\left( i \right)}_{\mu\nu} -2 \frac{\lambda_i}{n} \sum_{\mu\in \cal I_1,\nu\in \mathcal I_2} z_{\mu i}z_{\nu i}G^{\left( i \right)}_{\mu\nu},  \ \ \text{for }\  i \in \cal I_0 , \label{0self_Gii}\\
\frac{1}{{G_{\mu\mu} }}&=  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}\lambda_i \lambda_j z_{\mu i}z_{\mu j} G^{\left(\mu\right)}_{ij}, \quad \frac{1}{{G_{\nu\nu} }}=  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}  z_{\nu i}z_{\nu j}  G^{\left(\nu\right)}_{ij},  \ \ \text{for }\  \mu \in \cal I_1, \ \nu\in \cal I_2, \label{0self_Gmu1}
\end{align}
where we recall the notations in equation \eqref{labelZ}. For the right-hand side of equation \eqref{0self_Gii}, notice that the resolvent minor $G^{(i)}$ is defined such that it is independent of the entries $z_{\mu i}$ and $z_{\nu i}$. Hence by the concentration inequalities in Lemma \ref{largedeviation}, we have that the  right-hand side of equation \eqref{0self_Gii} concentrates around the partial expectation over the entries $\{z_{\mu i}: \mu \in \cal I_1\cup \cal I_2\}$, i.e., with high probability,
\begin{align*}
\frac{1}{{G_{ii} }}&=  - z - \frac{\lambda_i^2}{n} \sum_{\mu \in \mathcal I_1}  G^{\left( i \right)}_{\mu\mu} - \frac{1}{n} \sum_{\mu\in \mathcal I_2} G^{\left( i \right)}_{\mu\mu} +\oo(1)= - z - \lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1^{(i)}(z)-  \frac{\rho_2}{\rho_1+\rho_2} m_2^{(i)}(z)+\oo(1),
\end{align*}
where we used the definition of $m_1^{(i)}$ and $m_2^{(i)}$ in equation \eqref{defm} with $G$ replaced by $G^{(i)}$. Intuitively, since we have removed only one column and one row out of the $(p+n)$ columns and rows in $H$, $m_1^{(i)}$ and $m_2^{(i)}$ should be close to the original $m_1$ and $m_2$. Hence we obtain from the above equation that
\begin{align}\label{1self_Gii}
 G_{ii}  = -\left( z +\lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1(z) +  \frac{\rho_2}{\rho_1+\rho_2}m_2(z)+\oo(1)\right)^{-1}.
\end{align}
Similarly, we can obtain from equation \eqref{0self_Gmu1} that for $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\be\label{1self_Gmu} G_{\mu \mu }=-\left(1+\frac{p}{n} m_0 + \oo(1)\right)^{-1},\quad G_{\nu\nu}=-\left(1+\frac{p}{n} m+\oo(1)\right)^{-1},\ee
with high probability.

>>>>

\subsection{The Bias and Variance Asymptotics}
Finally, in this subsection, we describe briefly the main ideas in the proofs of Lemmas \ref{lem_cov_shift} and \ref{lem_cov_derivative}.
%We first describe the proof of Theorem \ref{lem_cov_shift_informal}.
%We briefly describe the key ideas for proving Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative}.
We use the standard Stieltjes transform method (or the resolvent method) in random matrix theory \cite{bai2009spectral,tao2012topics,erdos2017dynamical}. Both Lemma \ref{lem_cov_shift} and Lemma \ref{lem_cov_derivative} require a study of the matrix inverse $(X_1^\top X_1+X_2^\top X_2)^{-1}$ for $X_1= Z_1\Sigma_1^{1/2}$ and $X_2= Z_2\Sigma_2^{1/2}$.
\HZ{Suggestion: focus on the proof overview of Lemma \ref{lem_cov_shift} only. Use one paragraph in the end to say how to extend the idea to Lemma \ref{lem_cov_derivative}}
Suppose $M=\Sig_1^{1/2} \Sig_2^{-1/2}$ has a singular value decomposition
\be\label{eigen2}
M= U\Lambda V^\top, \quad \Lambda=\text{diag}( \lambda_1, \ldots, \lambda_p).
\ee
Then using equation \eqref{eigen2}, we can write
\be\label{eigen2extra}(X_1^\top X_1+X_2^\top X_2)^{-1}= \Sigma_2^{-1/2}V\left(   \Lambda U^\top Z_1^\top Z_1 U\Lambda  + V^\top Z_2^\top Z_2V\right)^{-1}V^\top\Sigma_2^{-1/2}.\ee
We divide the discussion into the following parts.

@@

 Using Schur complement formula (see {equation} (\ref{resolvent2})), we have that %for $i \in \cal I_0$, $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\HZ{This part needs more explanation - cannot understand.}
\begin{align}
\frac{1}{{G_{ii} }}&=  - z - \frac{\lambda_i^2}{n} \sum_{\mu,\nu\in \mathcal I_1} z_{\mu i}z_{\nu i}G^{\left( i \right)}_{\mu\nu} - \frac{1}{n} \sum_{\mu,\nu\in \mathcal I_2} z_{\mu i}z_{\nu i}G^{\left( i \right)}_{\mu\nu} -2 \frac{\lambda_i}{n} \sum_{\mu\in \cal I_1,\nu\in \mathcal I_2} z_{\mu i}z_{\nu i}G^{\left( i \right)}_{\mu\nu},  \ \ \text{for }\  i \in \cal I_0 , \label{0self_Gii}\\
\frac{1}{{G_{\mu\mu} }}&=  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}\lambda_i \lambda_j z_{\mu i}z_{\mu j} G^{\left(\mu\right)}_{ij}, \quad \frac{1}{{G_{\nu\nu} }}=  - 1 - \frac{1}{n} \sum_{i,j\in \mathcal I_0}  z_{\nu i}z_{\nu j}  G^{\left(\nu\right)}_{ij},  \ \ \text{for }\  \mu \in \cal I_1, \ \nu\in \cal I_2, \label{0self_Gmu1}
\end{align}
where we recall the notations in equation \eqref{labelZ}. For the right-hand side of equation \eqref{0self_Gii}, notice that the resolvent minor $G^{(i)}$ is defined such that it is independent of the entries $z_{\mu i}$ and $z_{\nu i}$. Hence by the concentration inequalities in Lemma \ref{largedeviation}, we have that the  right-hand side of equation \eqref{0self_Gii} concentrates around the partial expectation over the entries $\{z_{\mu i}: \mu \in \cal I_1\cup \cal I_2\}$, i.e., with high probability,
\begin{align*}
\frac{1}{{G_{ii} }}&=  - z - \frac{\lambda_i^2}{n} \sum_{\mu \in \mathcal I_1}  G^{\left( i \right)}_{\mu\mu} - \frac{1}{n} \sum_{\mu\in \mathcal I_2} G^{\left( i \right)}_{\mu\mu} +\oo(1)= - z - \lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1^{(i)}(z)-  \frac{\rho_2}{\rho_1+\rho_2} m_2^{(i)}(z)+\oo(1),
\end{align*}
where we used the definition of $m_1^{(i)}$ and $m_2^{(i)}$ in equation \eqref{defm} with $G$ replaced by $G^{(i)}$. Intuitively, since we have removed only one column and one row out of the $(p+n)$ columns and rows in $H$, $m_1^{(i)}$ and $m_2^{(i)}$ should be close to the original $m_1$ and $m_2$. Hence we obtain from the above equation that
\begin{align}\label{1self_Gii}
 G_{ii}  = -\left( z +\lambda_i^2 \frac{\rho_1}{\rho_1+\rho_2} m_1(z) +  \frac{\rho_2}{\rho_1+\rho_2}m_2(z)+\oo(1)\right)^{-1}.
\end{align}
Similarly, we can obtain from equation \eqref{0self_Gmu1} that for $\mu \in \cal I_1$ and $\nu\in \cal I_2$,
\be\label{1self_Gmu} G_{\mu \mu }=-\left(1+\frac{p}{n} m_0 + \oo(1)\right)^{-1},\quad G_{\nu\nu}=-\left(1+\frac{p}{n} m+\oo(1)\right)^{-1},\ee
with high probability.
